---
title: 'Futility of Utility Modelling'
author: Neil Stewart
#date: '`r format(Sys.time(), "%d %B %Y")`'
#bibliography: refs.bib
#csl: https://raw.githubusercontent.com/citation-style-language/styles/master/apa.csl
output: 
   html_document:
    toc: true
    toc_depth: 3 
---

---

```{r setup, echo=FALSE}
library(data.table)
library(ggplot2)
library(boot)
```

```{r}
sessionInfo()
```

This code generates Figures 5 and 6 in Stewart, Canic, and Mullett (2020) "On the futility of estimating utility functions: Why the parameters we measure are wrong, and why they do not generalize"

# Read in data

$px$ is the risky choice

```{r Load data and delete columns}
load("cleaned_data.RData")

data <- as.data.table(data)

# Delete unnecessary columns
data[, left.prob:=NULL]
data[, left.amount:=NULL]
data[, right.prob:=NULL]
data[, right.amount:=NULL]
data[, button:=NULL]
data[, RT:=NULL]
data[, location:=NULL]
data[, risky.choice:=NULL]
data[, prev.button:=NULL]
data[, alternation:=NULL]
data[, condition:=NULL] # It matched cond
data[, sub.no:=NULL] # We still have sub.id
data[, PA.id:=NULL]
data[, choice:=NULL] # We still have safe.choice dummy

str(data)
```


```{r Identify choice sets}
# Make sure we are working with integers for amounts
str(  positive.amounts <- as.integer(c(10, 20, 50, 100, 200, 500))  )
str(  uniform.amounts <- as.integer(c(100, 200, 300, 400, 500))  )

data[, choice.set.positive:= ifelse((x %in% positive.amounts) & (y %in% positive.amounts), 1L, 0L)]
data[, choice.set.uniform:= ifelse((x %in% uniform.amounts) & (y %in% uniform.amounts), 1L, 0L)]

data[cond=="Both" &  choice.set.positive &  choice.set.uniform, choice.set:="Common"]
data[cond=="Both" & !choice.set.positive & !choice.set.uniform, choice.set:="Unique"]
data[cond=="Both" & !choice.set.positive &  choice.set.uniform, choice.set:="Uniform"]
data[cond=="Both" &  choice.set.positive & !choice.set.uniform, choice.set:="Positive"]
```


```{r Check numbers of people and choices}
# 89 people in the Both Condition
unique(data[,.(sub.id, cond)])[,.N,by=.(cond)]

# 140 choices in the Both Condition; 150 in Positive and Uniform Conditions
unique(data[,.N, by=.(sub.id, cond)][order(cond),][,.(cond, N)])
```

# Figure 6 

```{r Figure 6}
log.odds <- function(p) { log(p/(1 - p)) }

#log.odds(0.01+0.98*safe.choice)

log.odds.safe <- data[cond=="Both", .(prop.safe.choice=mean(safe.choice)), by=.(log.y.over.x=log(y/x), choice.set, y, x)][order(log.y.over.x, choice.set),]
log.odds.safe[, log.odds.safe.choice:=log.odds(prop.safe.choice)]


retro.split <- rbind(
 	log.odds.safe[choice.set %in% c("Positive", "Common"),][,choice.subset:="Positive"],
 	log.odds.safe[choice.set %in% c("Uniform", "Common"),][,choice.subset:="Uniform"]
)

labels <- data.frame(x=c(-2.5, -0.78), y=c(0.13, 1.36), label=c("alpha==0.62 95\\% CI[0.54-0.70]", "alpha=0.94 95\\% CI[0.78-1.10]"))

labels <- data.frame(x=c(-3.2, -0.8), y=c(0.1, -0.6), label=c('atop(alpha[Positive]==0.62*",",95*"%"~CI~bgroup("[",0.54-"0.70","]"))', 'atop(alpha[Uniform]==0.94*",",95*"%"~CI~bgroup("[",0.78-"1.10","]"))'), choice.subset=c("Positive", "Uniform"))

x.axis.labels <- rbind(CJ(y=positive.amounts, x=positive.amounts), CJ(y=uniform.amounts, x=uniform.amounts))
x.axis.labels <- x.axis.labels[y<x,]
x.axis.labels <- unique(x.axis.labels)
x.axis.labels[,log.y.over.x:=log(y/x)]
x.axis.labels[,label:=paste("log(",y,"/",x,")",sep="")]
x.axis.labels <- aggregate(label ~ log.y.over.x, FUN=paste, data=x.axis.labels, collapse="\n")

ggplot(retro.split, aes(x=log.y.over.x, y=log.odds.safe.choice, col=choice.subset, shape=choice.subset)) + geom_point(size=2) + geom_smooth(method="lm", se=FALSE) + labs(col="Choice Set", shape="Choice Set", x=expression(log(y/x))) + geom_text(data=labels, mapping=aes(x=x, y=y, label=label,col=choice.subset, shape=NULL), parse=TRUE, show.legend=FALSE) + ylab(expression(log*bgroup("[",over(P(Safe),1-P(Safe)),"]"))) + scale_x_continuous(breaks=x.axis.labels$log.y.over.x, labels=x.axis.labels$label) + theme_classic() + theme(axis.text.x=element_text(angle=90, hjust=0, vjust=0.5)) + scale_color_brewer(palette="Dark2")

ggsave("log_odds_safe_by_log_y_over_x.pdf", width=9, height=6)#, scale=0.5)
```

# Run logistic regressions to fit free utility model to generate Figure 5

Fits to participants in the Both Condition, separately for the Positive and Uniform choice sets. Fits are of a non-parameteric model from Appendix A of Stewart, Reimers, and Harris (2015), and estimate the utility of each sum of money as a free parameter, rather than imposing a power-law functional form

```{r Add amount dummies}
# Add amount dummies

map <- read.csv("amount_map.csv")
map$cond <- factor(map$cond, levels=c("Positive", "Uniform", "Both"))
map$name <- paste("w", map$x, ".", map$cond, sep="")
map

data <- merge(data, map, by.x=c("x", "cond"), by.y=c("x", "cond"))
data <- merge(data, map, by.x=c("y", "cond"), by.y=c("x", "cond"), suffixes=c("x", "y"))

X.width <- max(map$i)
X <- matrix(0,nrow=nrow(data), ncol=X.width)
X[matrix(c(1:length(data$ix),data$ix),ncol=2)] <- -1
X[matrix(c(1:length(data$iy),data$iy),ncol=2)] <- 1
colnames(X) <- map$name[order(map$i)]

data <- cbind(data,X)
```


## Fit Positive and Uniform Choice Subsets from Both Condition

```{r Fit free model}
data.Both <- droplevels(subset(data,cond=="Both"))
amounts.Positive <- c(10,20,50,100,200,500)
amounts.Uniform <- c(100,200,300,400,500)

data.Both$Positive.subset <- with(data.Both, ifelse(x%in%amounts.Positive & y%in%amounts.Positive, TRUE, FALSE))
data.Both$Uniform.subset <- with(data.Both, ifelse(x%in%amounts.Uniform & y%in%amounts.Uniform, TRUE, FALSE))

fit.glm.nonpara.to.Both <- function(data, model) {
	switch(model, 
		Positive={m <- glm(safe.choice ~ log(q/p) + w10.Both + w20.Both + w50.Both + w100.Both + w200.Both, data=data, family=binomial)},
		Uniform={m <- glm(safe.choice ~ log(q/p) + w100.Both + w200.Both + w300.Both + w400.Both, data=data, family=binomial)},
		Both={m <- glm(safe.choice ~ log(q/p) + w10.Both + w20.Both + w50.Both + w100.Both + w200.Both + w300.Both + w400.Both, data=data, family=binomial)},
		stop("Uncaught condition in fit.glm.nonpara.to.Both()")
	)
	beta <- coef(m)	
	bias <- exp(beta["(Intercept)"])
	gamma <- beta["log(q/p)"]
	v <- exp( beta[grep("w", names(beta))] / gamma )
	data.frame(sub.id=data$sub.id[1], condition=data$cond[1], population=data$population[1], name=c("bias", "gamma", c(names(v),paste("w500", "Both", sep=".")), "logLik", "converged"), value=c(bias, gamma, v, 1, logLik(m), m$converged))
}

data.Both.Positive <- subset(data.Both, Positive.subset)
fits.m.nonpara.Both.Positive <- by(data.Both.Positive, INDICES=list(data.Both.Positive$sub.id), FUN=fit.glm.nonpara.to.Both, model="Positive")
fits.m.nonpara.Both.Positive <- do.call("rbind", fits.m.nonpara.Both.Positive)
fits.m.nonpara.Both.Positive <- merge(fits.m.nonpara.Both.Positive, map, all.x=TRUE)
fits.m.nonpara.Both.Positive <- fits.m.nonpara.Both.Positive[order(fits.m.nonpara.Both.Positive$sub.id),]
fits.m.nonpara.Both.Positive$subset <- "Positive"

data.Both.Uniform <- subset(data.Both, Uniform.subset)
fits.m.nonpara.Both.Uniform <- by(data.Both.Uniform, INDICES=list(data.Both.Uniform$sub.id), FUN=fit.glm.nonpara.to.Both, model="Uniform")
fits.m.nonpara.Both.Uniform <- do.call("rbind", fits.m.nonpara.Both.Uniform)
fits.m.nonpara.Both.Uniform <- merge(fits.m.nonpara.Both.Uniform, map, all.x=TRUE)
fits.m.nonpara.Both.Uniform <- fits.m.nonpara.Both.Uniform[order(fits.m.nonpara.Both.Uniform$sub.id),]
fits.m.nonpara.Both.Uniform$subset <- "Uniform"

fits.m.nonpara.Both <- Reduce(rbind, list(fits.m.nonpara.Both.Positive, fits.m.nonpara.Both.Uniform))

fits.m.nonpara.Both.molten <- melt(fits.m.nonpara.Both, id=c("sub.id", "condition", "population", "name", "subset"), measure.vars=c("value"))
fits.m.nonpara.Both.wide <- dcast(data=fits.m.nonpara.Both.molten, formula= ... ~ name + subset)

fits.m.nonpara.Both.wide <- fits.m.nonpara.Both.wide[,colSums(is.na(fits.m.nonpara.Both.wide))!=nrow(fits.m.nonpara.Both.wide)] # Drop all entirely NA columns

# At this point fits.m.nonpara.Both.wide is one row per participant, with a fit to only their positive choice set and a fit to only their uniform choice set
```

## Bootstrap medians

```{r Bootstrap medians}
boot.wrapper.m.nonpara <- function(data, indices, fun) {
	d <- data[indices,]
	apply(d[,grepl("^w", names(d)) & !grepl("^w500", names(d))], c(2), fun, na.rm=TRUE)
}

make.CI.table <- function(results) {
	estimate.and.CI <- function(bootci) {
		c(estimate=bootci$t0, CI95.=bootci$norm[2:3])
	}
	CIs <- vector("list", length(results$t0))
	for(i in 1:length(results$t0)) {
		CIs[[i]] <- boot.ci(results, type="norm", index=i)
	}
	tbl <- t(sapply(CIs, estimate.and.CI))
	colnames(tbl) <- c("estimate", "lower.CI", "upper.CI")
	rownames(tbl) <- names(results$t0)
	tbl
}

make.values.from.CI.table <- function(CI.table, map) {
	CI.table$name <- rownames(CI.table) # Drop everything after -
	CI.table$name <- sapply(row.names(CI.table), function(x) {strsplit(x, "_")[[1]][1]})
	CI.table$subset <- sapply(row.names(CI.table), function(x) {strsplit(x, "_")[[1]][2]})
	values <- merge(map, CI.table, all.x=TRUE)
	values[is.na(values)] <- 1
	values <- values[order(values$cond, values$x),]
	values
}

boot.result.median.nonpara.Both <- boot(data=fits.m.nonpara.Both.wide, statistic=boot.wrapper.m.nonpara, fun=median, R=10001, parallel="multicore", ncpus=23)

map.Both <- Reduce(rbind, list(cbind(subset(map, x %in% amounts.Positive & cond=="Both"), subset="Positive"),cbind(subset(map, x %in% amounts.Uniform & cond=="Both"), subset="Uniform")))

values.Both <- make.values.from.CI.table(as.data.frame(make.CI.table(boot.result.median.nonpara.Both)), map.Both)

write.csv(droplevels(subset(values.Both, subset!="All")), file="glm_nonpara_individual_bootstrap_both_data_points.csv")
```

## Figure 5

```{r Figure 5}
glm.nonpara.fit <- fread("glm_nonpara_individual_bootstrap_both_data_points.csv")

ggplot(glm.nonpara.fit, aes(x=x, y=estimate, col=subset, shape=subset, ymin=lower.CI, ymax=upper.CI)) + geom_line() + geom_linerange() + geom_point(size=3) + labs(x="Amount / £", y="Utility", col="Choice Subset", shape="Choice Subset") + theme_classic() + scale_color_brewer(palette = "Dark2") # + geom_line(data=power.functions.data, aes(x=x, y=utility, ymin=NULL, ymax=NULL, col=subset)) 

ggsave("glm_nonpara_individual_bootstrap_both.pdf", width=6, height=4)#, scale=0.5)
```


